---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r Load data}
library(ggplot2)
library(scales)

unzip(zipfile="activity.zip")
dataset <- read.csv('activity.csv')
str(dataset)

dataset$date <- as.Date(dataset$date)
summary(dataset)
```


## What is mean total number of steps taken per day?
```{r Mean Total steps per day}
### Calcuating total steps taken per day
tot_steps <- aggregate(dataset$steps, list(dataset$date), sum)
tot_steps <- as.data.frame(tot_steps)
names(tot_steps) <- c("Date", "Steps")

head(tot_steps)

### Making a histogram of total number of steps per day
ggplot(tot_steps, aes(x = Date, y = Steps)) + geom_bar(stat = 'identity') + scale_x_date(breaks = '1 day', limits = c(min(tot_steps$Date), max(tot_steps$Date)), labels = date_format("%d-%m"))+ theme_bw() + theme(axis.text.x = element_text(angle=90))

### Calculating the mean and median steps a day
mean_steps <- mean(tot_steps$Steps, na.rm = TRUE)
median_steps <- median(tot_steps$Steps, na.rm = TRUE)
```
__Mean : `r mean_steps`__
__Median : `r median_steps`__


## What is the average daily activity pattern?
```{r Daily Activity Pattern}
pattern <- aggregate(steps~interval, data = dataset, FUN = mean)
maxsteps <- max(pattern$steps)
maxinterval <- pattern[pattern$steps == maxsteps, "interval"]

ggplot(data = pattern, aes(x = interval, y=steps)) + geom_point() + labs(x = 'Time Interval', y='Avergae Steps', title = 'Average Daily Pattern') + theme_bw()
```
__The maximum average number of steps (`r maxsteps`) are taken in the interval `r maxinterval`.__


## Imputing missing values

We have `r sum(!complete.cases(dataset))` records with missing values. We have been suggested two ways to replace missing values

1. Replace missing step data for an interval with daily average for that interval
2. Replace missing step data for an interval with that day's average

Lets use the interval average over all days to replace a missing steps count for an interval. In other words, lets use method 1.
```{r Missing Values}
missing_vals <- is.na(dataset$steps)
missing_intervals <- dataset$interval[missing_vals]
input_steps <- pattern$steps[match(missing_intervals, pattern$interval)]
new_dataset <- dataset
new_dataset$steps[missing_vals] <- input_steps
```

Lets recalculate the mean and median with the corrected data.
```{r Corrected Data}
corrected_dailysteps <- aggregate(steps~date, new_dataset, sum)
new_mean <- mean(corrected_dailysteps$steps)
new_median <- median(corrected_dailysteps$steps)
ggplot(corrected_dailysteps, aes(x = date, y = steps)) + geom_bar(stat = 'identity') + scale_x_date(breaks = '1 day', limits = c(min(corrected_dailysteps$date), max(corrected_dailysteps$date)), labels = date_format("%d-%m"))+ theme_bw() + theme(axis.text.x = element_text(angle=90))
```
__Mean : `r new_mean`__

__Median : `r new_median`__

## Are there differences in activity patterns between weekdays and weekends?

```{r Weekdays vs Weekends}
new_dataset$day_type = "Weekday"
weekend <- c("Saturday", "Sunday")
new_dataset$day_type[weekdays(new_dataset$date) %in% weekend] <- "Weekend"
new_dataset$day_type <- as.factor(new_dataset$day_type)

corrected_activitypattern <- aggregate(steps~interval+day_type, new_dataset, mean)

ggplot(data = corrected_activitypattern, aes(x = interval, y=steps)) + geom_point() + geom_line() + facet_grid(day_type ~ .)
```